# Creation of the snow_depth dataframe containing the daily snow depth in cm for countries in Europe
# Link to data:  https://www.ecad.eu/dailydata/predefinedseries.php  -  "Daily snow depth SD" - ECA_blend_sd.zip

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(purrr)
```

# Create the dataframe 
```{r}
# Create a list of all text files starting with "SD_"
files <- list.files(pattern = "^SD_.*\\.txt$")

# Read the files and skip the first 23 lines 
snow_depth <- lapply(files, function(file) {
  data <- read.table(file, header = FALSE, sep = ",", skip = 23, fill = TRUE)
  return(data)
})

# Combine the files into a dataframe
snow_depth <- do.call(rbind, snow_depth)
```

# Delete and rename columns
```{r}
# Delete the unnecessary SOUID: Source identifier and Q_SD quality check columns
snow_depth <- snow_depth |> select(-V2, -V5)

# Rename the columns
snow_depth <- snow_depth |> rename(id = V1, date = V3, snow_depth = V4)
```

# Create a dataframe from the stations.txt file containing additional information for each station
```{r}
snow_depth_stations <- read.delim(file = "stations.txt", header = FALSE, skip = 19, sep = ",")

# Rename the columns
snow_depth_stations <- snow_depth_stations |> rename(id = V1, station = V2, country = V3, latitude = V4, longitude = V5, height = V6)
```

# Merge the snow_depth_stations and snow_depth dataframes
```{r}
snow_depth <- merge(snow_depth_stations, snow_depth, by = "id", all = FALSE)

# Place the date column after the id column and the station column after the country column
snow_depth <- snow_depth |> 
  relocate(date, .after = id) |> 
  relocate(station, .after = country)
```

# Clean and convert data
```{r}
# Remove leading and trailing whitespace from the country and station columns
snow_depth <- snow_depth |>
  mutate(country = trimws(country),
         station = trimws(station))

# Rename the ISO 3166 country codes to full country names
snow_depth <- snow_depth |>
  mutate(country = recode(country, 
                           "AM" = "armenia",
                           "AT" = "austria",
                           "AZ" = "azerbaijan",
                           "BY" = "belarus",
                           "CH" = "switzerland",
                           "DE" = "germany",
                           "DK" = "denmark",
                           "EE" = "estonia",
                           "ES" = "spain",
                           "FI" = "finland",
                           "FR" = "france",
                           "GB" = "united kingdom",
                           "GE" = "georgia",
                           "GL" = "greenland",
                           "HR" = "croatia",
                           "HU" = "hungary",
                           "IE" = "ireland",
                           "IS" = "iceland",
                           "IT" = "italy",
                           "KZ" = "kazakhstan",
                           "KG" = "kyrgyzstan",
                           "LV" = "latvia",
                           "LT" = "lithuania",
                           "LU" = "luxembourg",
                           "MD" = "moldova",
                           "ME" = "montenegro",
                           "NL" = "netherlands",
                           "NO" = "norway",
                           "PL" = "poland",
                           "RS" = "serbia",
                           "RU" = "russia",
                           "SE" = "sweden",
                           "SI" = "slovenia",
                           "SK" = "slovakia",
                           "SJ" = "svalbard and jan mayen",
                           "TM" = "turkmenistan",
                           "TJ" = "tajikistan",
                           "UA" = "ukraine",
                           "UZ" = "uzbekistan"))


# Convert all station names to lower case
snow_depth$station <- tolower(snow_depth$station)

# Convert the numeric format of the "date" column to the date format 
snow_depth$date <- as.Date(as.character(snow_depth$date), format = "%Y%m%d", origin = "1970-01-01")

# Convert values of -9999 to NAs
snow_depth[snow_depth == -9999] <- NA

# Convert the int data type of the snow_depth column to num so it matches the one in the dataframe with the missing stations
snow_depth$snow_depth <- as.numeric(snow_depth$snow_depth)

# Drop all NAs in the dataframe
# snow_depth <- na.omit(snow_depth)
```

