# Merge the temperature and precipitation dataframes and calculate snow_depth data for the countries Finland, France, Italy and Spain
```{r}
library(tidyverse)
library(dplyr)
```

# Merge the temperature and precipitation dataframes 
```{r}
countries <- merge(temperature, precipitation, by = c("id", "station", "country", "latitude", "longitude", "height", "date"), all = FALSE)

# Place the date column after the id column and the station column after the country column
countries <- countries |> 
  relocate(date, .after = id) |> 
  relocate(station, .after = country)

# Convert precipitation values from 0.1 mm to 1 mm
countries$precipitation <- countries$precipitation / 10

# Convert temperature values from 0.1 Degrees Celsius to 1 Degrees Celsius
countries$temperature <- countries$temperature / 10
```


# Create the snow_depth column and calculate the snow depth
```{r}
# Sort the dataframe by country, then by station, then by date
countries <- countries |> arrange(country, station, date)

# Create the snow_depth column with all values set to 0
countries$snow_depth <- 0

# Initialize previous snow depth value
prev_snow_depth <- 0

# Loop through each row
for (i in 1:nrow(countries)) {
  # If temperature is less than 0
  if (countries$temperature[i] < 0) {
    # Add precipitation to snow depth
    countries$snow_depth[i] <- countries$snow_depth[i] + countries$precipitation[i]
  }
  
  # Calculate 10% of snow_depth of the previous day
  tenth <- 0.1 * prev_snow_depth
  
  # Calculate the amount of snow to transfer from the previous day according to current temperature and snow quantity of the previous day  
  if (countries$temperature[i] < 5) {
    if (tenth < 1) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (max(0, prev_snow_depth - 1))
    } else if (tenth > 5) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (prev_snow_depth - 5)
    } else {
      countries$snow_depth[i] <- countries$snow_depth[i] + (0.9 * prev_snow_depth)
    }
  } else if (countries$temperature[i] >= 5 && countries$temperature[i] < 10) {
    if (tenth < 2) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (max(0, prev_snow_depth - 2))
    } else if (tenth > 8) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (prev_snow_depth - 8)
    } else {
      countries$snow_depth[i] <- countries$snow_depth[i] + (0.8 * prev_snow_depth)
    }
  } else {
    if (tenth < 3) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (max(0, prev_snow_depth - 3))
    } else if (tenth > 10) {
      countries$snow_depth[i] <- countries$snow_depth[i] + (prev_snow_depth - 10)
    } else {
      countries$snow_depth[i] <- countries$snow_depth[i] + (0.5 * prev_snow_depth)
    }
  }
  
  # If the temperature is 2 or more, precipitation is rain and melts away snow at a ratio of precipitation / 2
  if (countries$temperature[i] >= 2) {
    countries$snow_depth[i] <- countries$snow_depth[i] - (countries$precipitation[i] / 2)
  }
  
  # Round snow depth to 1 decimal place and ensure it's neither negative nor higher than 700
  countries$snow_depth[i] <- max(min(round(countries$snow_depth[i], 1), 700), 0)
  
  # Set previous snow depth to current snow depth for next iteration
  prev_snow_depth <- countries$snow_depth[i]
}

```

# Delete the temperature and precipitation columns 
```{r}
countries <- countries |> select(-temperature, -precipitation)
```

# Combine the countries dataframe with the snow_depth dataframe to create the final dataframe containing snow_depth data for all countries in Europe
```{r}
cdk1 <- rbind(snow_depth, countries)

# Sort the dataframe by country, then by station, then by date
cdk1 <- cdk1 |> arrange(country, station, date)
```

