---
title: "R Notebook"
output: html_notebook
---
```{r}

library(tidyverse)
library(lubridate)
```

```{r}
load("poland.Rdata")
# load("switzerland.Rdata")
# load("slovakia.Rdata")

# save(poland, file = "poland.Rdata")

# str(poland)
# unique(poland$staname)
# unique(poland$stahght)

```

# Create a column with Station name and height combined

```{r}
# poland <- poland %>%
#   mutate(stahght = paste0(stringr::str_to_title(staname), ", ", hght, "m"))

# Show a list of stations arranged by their height

# stations_by_height <- poland %>% 
#   select(staname, hght) %>% 
#   distinct() %>% 
#   arrange(hght)

```

# Plot of extreme events - The largest and smallest amounts of snow for the winter months during a specified time period

```{r}

winter_months <- poland %>%
  filter(stahght == "Hala Gasienicowa, 1379m") %>%
  filter(year(date) >= 1950 & year(date) <= 2023) %>%
  mutate(month = lubridate::month(date),
         month_label = case_when(
           month == 11 ~ "Nov",
           month == 12 ~ "Dec",
           month == 1 ~ "Jan",
           month == 2 ~ "Feb",
           month == 3 ~ "Mar",
           month == 4 ~ "Apr"
         )) %>%
  filter(month %in% c(11, 12, 1, 2, 3, 4)) %>%
  group_by(year = lubridate::year(date), month_label) %>%
  summarise(mean_sd = mean(sd)) %>%
  mutate(month_label = factor(month_label, levels = c("Nov", "Dec", "Jan", "Feb", "Mar", "Apr"), ordered = TRUE))



combined_obs <- bind_rows(
  winter_months %>%
    group_by(month_label) %>%
    filter(mean_sd == min(mean_sd)) %>%
    arrange(desc(year)) %>%
    slice(1) %>%
    transform(obs_type = "Smallest"),
  winter_months %>%
    group_by(month_label) %>%
    slice_max(mean_sd) %>%
    transform(obs_type = "Largest")
)



ggplot(combined_obs, aes(x = month_label, y = mean_sd, fill = obs_type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = year), position = position_dodge(width = 0.9), vjust = -0.5, size = 4, color = "black") +
  labs(x = NULL, y = "Mean Snow Depth in cm", title = "Winter Months with Largest and Smallest Amounts of Snow",
       subtitle = "Years:      1950 - 2023\nLocation:  Hala Gasienicowa, 1379m") +
  facet_wrap(~ obs_type, ncol = 2, strip.position = "top") +
  scale_fill_manual(values = c("#2671B5", "#E9333F")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.10))) +  # Adjust the expand argument to add space in the y-axis
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "none",
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(10, 20, 20, 10),
    strip.text = element_blank(),
    strip.background = element_blank()
  )



```


# Line Plot: Snow Depth over the years with dashed line for reference value

```{r}

# Calculate the baseline of the average snow depth for years between 1920 and 1950 and winter months November until April

reference <- poland %>%
  filter(stahght == "Warszawa-Okecie, 107m") %>%
  filter(year(date) >= 1950 & year(date) <= 1980 & month(date) %in% c(11, 12, 1, 2, 3, 4)) %>%
  summarise(mean_sd = mean(sd)) %>%
  pull(mean_sd)


# Show a line plot



poland %>%
  filter(stahght == "Warszawa-Okecie, 107m") %>%
  filter(year(date) >= 1800 & month(date) %in% c(11, 12, 1, 2, 3, 4)) %>%
  group_by(year = (year(date))) %>%
  summarise(mean_sd = mean(sd)) %>%
  ggplot(aes(x = year, y = mean_sd)) +
  geom_smooth(se = FALSE, color = "#2671B5", linetype = "solid", size = 1.2) +
  geom_hline(yintercept = reference, color = "red", linetype = "dashed", size = 1.2) +
  labs(y = "Snow Depth in cm", x = "") +
  ggtitle("Mean Yearly Snow Depth from November to April") +
  labs(subtitle = "Dashed Line: Mean Snow Depth between 1950 and 1980") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    legend.position = "none",
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(10, 20, 00, 10),
    plot.subtitle = element_text(color = "red"),
    plot.caption = element_text(size = 14, face = "italic", margin = margin(0, 0, 15, 0))  # Adjust the font size, style, and margin of the caption
  ) +
  expand_limits(y = 0) +
  labs(caption = "Warszawa-Okecie (107m), Poland")




```


# Comparison between 3 different stations 

```{r}


poland %>%
  filter(stahght %in% c("Plock, 87m", "Zabkowice, 319m", "Sloszow, 636m")) %>%
  ggplot(aes(x = year(date), y = sd, color = stahght)) +
  geom_smooth(se = FALSE, size = 1.2) +
  labs(
    title = "Snow Depth of Selected Locations in Poland",
    x = "",
    y = "Mean Snow Depth in cm"
  ) +
  scale_color_manual(
    values = c("Plock, 87m" = "#E41A1C", "Zabkowice, 319m" = "#377EB8", "Sloszow, 636m" = "#66C2A5"),
    name = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Set hjust to center the title
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.box = "horizontal",
    legend.position = "top",
    legend.margin = margin(b = 10),
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  )



```



# Bar Plot for monthly comparison of two periods in in Sniezka, Poland

```{r}
monthly_avg_combined <- rbind(
  poland %>%
    filter(stahght == "Sniezka, 958m") %>%
    filter(year(date) >= 1950 & year(date) <= 1980) %>%
    group_by(month = factor(format(date, "%m"), levels = sprintf("%02d", 1:12), labels = month.abb)) %>%
    summarise(sd = mean(sd)) %>%
    mutate(dataset = "Monthly Average 1"),
  
  poland %>%
    filter(stahght == "Sniezka, 958m") %>%
    filter(year(date) >= 2018 & year(date) <= 2023) %>%
    group_by(month = factor(format(date, "%m"), levels = sprintf("%02d", 1:12), labels = month.abb)) %>%
    summarise(sd = mean(sd)) %>%
    mutate(dataset = "Monthly Average 2")
)



ggplot(monthly_avg_combined, aes(x = month, y = sd, fill = dataset)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7, color = "black", size = 0.5) +
  labs(title = "Comparison of Monthly Snow Depth between 2 Periods",
       subtitle = "Sniezka (958m), Poland",  # Add the subtitle
       x = "",  # Remove the "Month" label
       y = "Mean Snow Depth in cm") +
  scale_fill_manual(values = c("Monthly Average 1" = "#2671B5", "Monthly Average 2" = "#E9333F"), 
                    labels = c("1950 - 1980", "2018 - 2023")) +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 14, face = "italic", margin = margin(b = 10)),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.position = "top",
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())



```

# Plot snow depth by height range (this is of limited usability as there are more stations in later years, especially at higher altitudes, which skew the results)

```{r}


poland %>%
  filter(year(date) >= 1920) %>%
  group_by(year(date)) %>%
  mutate(height_range = case_when(
    hght < 450 ~ "Below 450m",
    hght >= 451 & hght <= 600 ~ "451m - 600m",
    hght > 600 ~ "Above 600m"
  )) %>%
  ggplot(aes(x = year(date), y = sd, color = height_range)) +
  geom_smooth(se = FALSE, size = 1.2) +
  labs(
    title = "Snow Depth of Various Height Ranges in Poland",
    x = "",  # Remove the "Year" label
    y = "Mean Snow Depth in cm",
    color = "Height Range"
  ) +
  scale_color_manual(values = c("Below 450m" = "#E41A1C", "451m - 600m" = "#377EB8", "Above 600m" = "#66C2A5")) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    legend.position = "right",
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank()
  )

```


# Plot showing mean station height measurements over the years

```{r}
library(ggplot2)
library(dplyr)

switzerland %>%
  group_by(year = lubridate::year(date)) %>%
  summarise(mean_hght = mean(hght)) %>%
  ggplot(aes(x = year, y = mean_hght)) +
  geom_line(color = "#2671B5", size = 1.2) +
  labs(y = "Meters above Sea Level", x = "Year") +
  ggtitle("Mean Height of measurements in Switzerland by Year") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    legend.position = "none",
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(10, 20, 20, 10)
  )

poland %>%
  group_by(year = lubridate::year(date)) %>%
  summarise(mean_hght = mean(hght)) %>%
  ggplot(aes(x = year, y = mean_hght)) +
  geom_line(color = "#2671B5", size = 1.2) +
  labs(y = "Meters above Sea Level", x = "Year") +
  ggtitle("Mean Height of measurements in Poland by Year") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    legend.position = "none",
    panel.grid.major = element_line(color = "lightgray"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    plot.margin = margin(10, 20, 20, 10)
  )

```




